Consistency scores in text data

### 2020-10-18 outputs/documents/paper_bert.Rmd
Pytorch Bert text-infilling

We consider various natural language processing methods. These include n-grams, Word2Vec, GloVe, GPT models and BERT. The n-gram approach involves identifying two, three, or more, words that are commonly found together. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance, an equally reasonable two-gram involving the word 'good' is 'good work'. For that reason, we consider pre-trained word embedding models. These include Word2Vec and GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship. We also consider pre-trained generative models such as GPT-2 and GPT-3 and BERT. GPT-2, GPT-3 and BERT are pre-trained generative unsupervised language models. GPT-2 and GPT-3 are the two generations of the same model that differ in the number of parameters. GPT-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters. While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API. BERT has 340 millions parameters which is smaller than both GPT-2 and GPT-3. BERT also differs from GPT models because its encoding process is bidirectional instead of unidirectional.

@@

## PDF text extraction and n-grams

The Portable Document Format (PDF) is typically used to render documents in a relatively fixed way. PDFs contain a lot of information, but often that information is not able to be analysed directly by statistical language models. The information first needs to be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared, and made into some type of dataset. One issue in this process is that OCR is not perfect. Hence there is a need to correct the output before it is analysed. One way to do this is to correct errors one-by-one and independent of context. We all know that 'tbe' is wrong and should be 'the'. But statistical language models have the advantage of correction that can occur within context (@rosenfeld2000two). This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat'.

The foundation of an n-gram language model is the conditional probability set-up introduced above. An n-gram model is a probabilistic language model that predicts the next word in a sequence of words (@bengio2003neural). The $n$ in n-gram refers to the number of words in that sequence. Consider the following excerpt from *Jane Eyre*: 'We had been wandering'. 'We' is a uni-gram, 'We had' is a bi-gram, 'We had been' is a tri-gram, and 'We had been wandering' is a 4-gram. Notice that the two tri-grams in this excerpt, 'We had been', and 'had been wandering', overlap. The use of n-gram models enables us to assign probabilities to both the next sequence of words and just the next word. For instance, consider the two sentences: 'We had been wandering', and 'We had been wangling'. The former is likely to be more frequently encountered in a training corpus. Thus, an n-gram would assign a higher probability to the next word being 'wandering' than 'wangling', given the sequence 'We had been'.

To predict the next word, we have to take the sequence of preceding words into account, which requires knowing the probability of the sequence of words. The probability of a sequence appearing in a corpus follows the chain rule:

$$\mbox{Pr(We,had,been,wandering)} = \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|We,had)} \\  
\times \mbox{Pr(wandering|We,had,been)}.$$

However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making prediction difficult. Alternatively, we can approximate the probability of a word depending on only the previous word. This is known as the 'Markov assumption' and it allows us to approximate the probability using only the last $n$ words [@brown1992class]:

$$\mbox{Pr(We,had,been,wandering)} \approx \mbox{Pr(We)} \times \mbox{Pr(had|We)} \times \mbox{Pr(been|had)} \\ 
\times \mbox{Pr(wandering|been)}.$$

As a bi-gram model only considers the immediately preceding word, under the Markov assumption, an $n$-gram model can be reduced to a bi-gram model with $n$ being any number:

$$\mbox{Pr}(w_{n} | w^{n-1}_{1}) \approx \mbox{Pr}(w_{n} | w_{n -1})$$

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation [@brown1992class]. In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability.

However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. 1] discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited.

There are a variety of ways to implement an n-gram model within R @citeR including using R packages such as `Quanteda` [@citequanteda], `tidyText` [@citetidytext] and `tm` [@citetm]. We used the Text Analysis Utilities, `tau`, R package @citetau because of we can separate text-cleaning and n-gram making in two different steps. We want to have control to manually clean the text because we want to add an end-of-sentence token to each sentence int he corpus. To illustrate the use of `tau`, we will first install and load the library, and create a small corpus.

```{r, eval = FALSE}
install.packages("tau")
```

```{r, eval=FALSE}
library(tau)
corpus <- "There was no possibility of taking a walk that day. We had been 
wandering, indeed, in the leafless shrubbery an hour in the morning; but 
since dinner (Mrs. Reed, when there was no company, dined early) the cold 
winter wind had brought with it clouds so sombre, and a rain so penetrating, 
that further out-door exercise was now out of the question."
```

Our purpose is to construct tri-grams and so we first split the corpus into sentences by period, exponential mark, or question marks.

First, we remove all punctuation and other non-letter characters, and convert all text to lower case.

```{r, eval=FALSE}
tokenizer <- function(corpus) {
  lines <- vector()
  # Read the corpus line by line
  for (line in corpus) {
    # add two start-of-sentence tokens to the beginning of a sentence
    lines <-c(lines, line)
  }
  lines <- tolower(lines)
  # add end-of-sentence token to the end of a sentence
  lines <- gsub("[.!?]$|[.!?] |$", " </s>", lines) 
  tokens <- unlist(strsplit(lines, "[^a-z<>/]"))
  tokens <- tokens[tokens != ""]
  return(tokens)
}
```

We will use the `textcnt()` function to turn the training corpus into tri-grams. The most important argument of this function is _corpus_, as well as _n_ to set the number of words in a sequence, _method_ to determine which type of counts to compute, and _split_ to designate the regular expression pattern for splitting. The function outputs a `textcnt` object, which is essentially a table with each tri-gram and their counts [@citetau].

```{r, eval=FALSE}
trigrams <- textcnt( tokens, n=3, method = "string", 
                     split = "[[:space:][:digit:]]+", decreasing = TRUE)
```

We then manually added an end of sentence token, `</s>`, to the end of each sentence.

```{r, eval=FALSE}
trigrams <- trigrams[!grepl("</s>", names(trigrams))]
```

At this point it is possible to take other measures around fine-tuning the tri-grams. For instance we may like to remove the tri-grams that contain an end-of-sentence token. This would mean that tri-grams that include a word from a new sentence will be removed. For example, the tri-gram 'day </s> we' will be removed because it has words from two different sentences. Table 1. shows four sample tri-grams from the small sample corpus after this process.

```{r table1, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| there was no  | a rain so | a walk that | an hour in  |
|---------------|-----------|-------------|-------------|
| 2             | 1         | 1           | 1           |

Table 1. Four sample tri-grams generated from the corpus
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```

The tri-grams can be accessed by calling the names of this table object and the counts can be used to calculate probabilities. There are a variety of options around constructing these probabilities. One is Kneser–Ney smoothing, which is the smoothing technique recommended by @chen1999empirical to improve the accuracy of n-gram models.

Finally, it is helpful to create a data table that contains the first two words and the last of each tri-gram in separate column and the smoothed probability in another column. For instance, Table 2. provides and example of this set-up.

```{r table2, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| First Words   | Last Word | Probability |
|---------------|-----------|-------------|
| there was     | no        | 0.6278947   |

## Word embedding models

Although, from an implementation perspective, there is an inherent numerical representation of language in n-gram models, it is not overly tractable. The numeric representation typically relies on each term being represented in a vector, which can quickly result in quite high-dimensional objects. In contrast, word embedding models represent terms by vectors of real numbers where each vector is a 'feature', hence the final object has considerably lower dimensionality [@bengio2003neural].

This approach has a long history, but the implementation of @bengio2003neural has become the foundation for much subsequent work. Here, terms are encoded into feature vectors that represent different aspects of the term, and so each term is represented by a position in the vector space. Terms that are used in similar ways will have similar representations in the vector space. This will be due to their co-occurrence in usage. For example, 'cat' will be closer to 'dog' than to 'car', because we are more likely to talk about cats and dogs, than cats and cars. See how they might be close when considered along a 'pets' dimension, but far when considered along a 'related-to-tigers' dimension.

Terms that are often used in similar contexts will have similar vector representations. Hence, word embedding models can identify similar terms, calculate the similarity between a pair of terms, and find the odd term in a group of terms [@mikolov2013efficient]. There are a variety of ways to achieve this dimensionality reduction, including Word2Vec and GloVe.

### Word2vec

Word2vec is a word embedding model that uses a neural network to learn word associations [@mikolov2013efficient] [@mikolov2013distributed]. It takes a corpus as an input and outputs a set of feature vectors that allow the terms in that corpus to be represented. @mikolov2013efficient proposed two architectures for determining this distributed representation of terms: continuous bag-of-words (CBOW), and continuous skip-gram. Both have an input layer, a projection layer, and an output layer. But CBOW uses several surrounding words, or the context, to predict a target word. On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words.

To illustrate the difference, consider the quote from Descartes, 'I think therefore I am'. The focus word for which we want to learn the embedding is 'therefore', and the context window is set to two. This means that we consider the two words immediately preceding and proceeding 'therefore'. The CBOW input is 'I, think, I, am', and the output is 'therefore'. In contrast, the continuous skip-gram input is 'therefore', while 'I, think, I, am' is the output. The input layer encodes terms within a context window to one-hot vectors. Using a projection matrix, the projection layer then project the discrete one-hot vectors to lower-dimension continuous vectors. Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words [@rong2014word2vec]. It is the input and output vectors that differ between the two architectures.

Word2Vec is trained on the Google News dataset [@citegooglenews]. This contains six billion tokens and the vocabulary size is restricted to the one million most common words [@mikolov2013efficient]. As this is a large dataset, it provides an accurate statistical pattern of language, as it is used in this dataset, and a comprehensive mapping between terms and vector positions.

To the extent that the relative positions of terms reflects their semantic relationship, Word2Vec will perform well for word analogies. However, terms that have context-specific meaning may not perform as well.

There are a few packages for implementing a word2vec model in R, including `rword2vec` [@citerwordvecchaware] and `word2vec` [@citewordvecwijffels]. Both packages use the same algorithm. Another way to implement a word2vec model in R is through Keras which is an API for deep learning written in Python [@chollet2015keras] for which the `keras` package provides an R interface [@citekeras].

Here, we follow @falbel2017word to implement a skip-gram model of Word2Vec through Keras. There are three steps in the Keras deep learning framework: 1) prepare the data; 2) define the model; and 3) train and evaluate the model. **(Ke-Li: Can we change this to be a different dataset please? Yes, I changed it to Google News for now, let's change it to an official data set that we will decide on later!)** Following these steps, we first downloaded the Google News dataset [@citegooglenews] and preprocess the text into tokens.

```{r, eval=FALSE}
tokenizer %>% fit_text_tokenizer(training_corpus)
```

We then define a model that mimics the skip-gram algorithm of word2vec using the function `skipgrams_generator()`.

```{r, eval=FALSE}
skipgrams_generator <- function(text, tokenizer, window_size, negative_samples) {
  gen <- texts_to_sequences_generator(tokenizer, sample(text))
  function() {
    skip <- generator_next(gen) %>%
      skipgrams(
        vocabulary_size = tokenizer$num_words, 
        window_size = window_size, 
        negative_samples = 1
      )
    x <- transpose(skip$couples) %>% map(. %>% unlist %>% as.matrix(ncol = 1))
    y <- skip$labels %>% as.matrix(ncol = 1)
    list(x, y)
  }
}
```

The embedding dimensions, the size of the context window and the number of samples are configurable within this function.

Finally, we use the `fit_generator()` function to train the model.

```{r, eval=FALSE}
model %>%
  fit_generator(
    skipgrams_generator(training_corpus, tokenizer, skip_window, negative_samples), 
    steps_per_epoch = 10, epochs = 1
  )
```

**What's supposed to happen:** As a result, a embedding matrix is obtained and serves as the look-up table of the word vectors. **What really happened:** However, the training process was stuck in the operation of `fit_generator()` function. According to Stack Flow this might be an issue of incompatible versions between Keras and Tensorflow. More investigation is needed.

### GloVe

The Global Vectors (GloVe) model is a pre-trained word embedding model that emphasises the semantic meaning of words [@pennington2014glove]. In the Word2Vec model, word representations are learnt locally within the context window. On the other hand, GloVe considers the properties of the corpus globally. As the model considers the context of an entire corpus the co-occurrence probabilities are based on counts for the entire training dataset.

GloVe additionally considers the relative probability of word co-occurrence. For instance, consider the two words 'ice' and 'steam'. GloVe considers their co-occurrence with a variety of probe words, $k$. For instance $k$ could be 'solid', which is a property of ice; 'gas', which is a property of steam; and 'water', which is related to both.The ratio of co-occurrence probability is:

$$\frac{\mbox{Pr}(k|\mbox{ice})}{\mbox{Pr}(k|\mbox{steam})}.$$

The ratio between the co-occurrence probability of the 'ice, solid' pair and the 'steam, solid' pair is high because the co-occurrence of 'ice' and 'solid' is higher than the co-occurrence of 'steam' and 'solid'. Conversely, the ratio between the co-occurrence of the 'ice, gas' pair and the 'steam, gas' pair is low. On the other hand, the ratio of the co-occurrence probability of the 'ice, water' pair and the 'steam, water' pair us close to one because water is relevant to both 'ice' and 'steam'.

As @hanretty2018comparing point out, although word embedding models can capture semantic meanings of words, "they are context-free and fail to capture higher-level concepts in context, such as polysemous disambiguation, syntactic structures, semantic roles, anaphora". Word embedding models do not represent a homonyms that has difference meanings in different vectors depending on the context. For example, "squash" can be a type of hard shelled vegetable or can be referring to a racket and ball sport, but there is only one series of vectors to represent the word "squash", instead of dynamically changing the vectors depending on the context. Moreover, from our observation, the word embedding models are unable to handle novel words that are not included in the vocabulary of the training corpus; for this reason, we conclude that the generalizability of word embedding models is sub-optimal.

There are a variety of ways to implement GloVe word embedding models in R. One is to load GloVe embeddings into a frozen Keras Embedding layer [@kerasglove] using the `Keras` R package [@kerasR2020]. This means the embedding vectors will not be updated during the training process. Another way, is to use the `text2vec` package [@text2vec].

Here we follow the blog post published by Keras [@kerasglove] to implement GloVe using the `keras` R package [@kerasglove]. The blog post is aimed to create a word embedding model for text classification. However, we will only use the guide create a embedding index and skip the steps of text classification. Instead of building a model for word embedding from scratch as we did for the skip-gram model of Word2Vec, here we are merely loading the Glove model and its pre-trained word vectors as an index. The word embedding index is then saved to an R environment.

To use GloVe in R, we first download the pre-trained GloVe vectors. The file comes in zip format, which needs to be unzipped:

```{r, eval=FALSE}
download.file('http://nlp.stanford.edu/data/glove.6B.zip',destfile = 'glove.6B.zip')
unzip('glove.6B.zip')
```

The next step is to load the GloVe vectors into R:

```{r, eval=FALSE}
vectors = data.table::fread('glove.6B.300d.txt', data.table = F,  encoding = 'UTF-8')
colnames(vectors) = c('word',paste('dim',1:300,sep = '_'))
```

The GloVe vectors are now available in R. We restructure the word vectors into a tibble data frame for easy access. The tibble dataset has 400,000 words and 301 dimensions to represent each word:

```{r, eval=FALSE}
as_tibble(vectors)
# A tibble: 400,000 x 301
```

@@

# Stylized example

**(TODO: This stylized example need to be changed and updated to actually use GPT-3 at some point.)**

In our application, we deploy OpenAI GPT-3 to generate text and ...

As a stylized example, let's consider the following actual paragraph from Jane Eyre, by Charlotte Bronte:

> There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs. Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.

Let's pretend that this text had been created from optical character recognition and that it had the following errors: some 'h' were replaced with 'b'; and some 'd' have been replaced with 'al':

> There was no possibility of taking a walk that day. We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs. Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question.

Assume a model that is trained to perfectly forecast the next word in Jane Eyre. For this fragment there are 62 words, comprising 5 errors and 57 correct words. So the internal consistency score of this fragment would be: 57/62 = 0.919. When we recognise and correct the errors, this consistency score would increase to 1.

Similarly, assume a model that is trained on an external data source. This means that it will recognise the the cases where some 'h' were replaced with 'b', but not recognise that 'Reed' has become 'Reeal'. Hence, the external consistency score would be 58/62 = 0.935.

# Data

Various data can be used....

# Application

# Discussion

Internal validity vs external- one is their own words the other is a general set of words.

In the same way that precision and recall provide important measures…

@@

## Literature summaries

(These are just here for now - we'll remove later.)

### Article: Improving Language Understanding by Generative Pre-Training (2018)
Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.

### Article: Language Models are Few-Shot Learners (2020)
Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.

## Discussion of how GPT-3 works

**(Ke-Li - it would be good for us to go through how it actually works in as much detail as we can manage (in our own words). This is mostly just to force us to learn how it works and may be taken out of the final version of the paper, but it's still an important exercise.)**

>>>>

We consider various Transformer language models. These include BERT [@devlin2018bert], GPT-2 [@radford2019language] and XLNet [@yang2019xlnet]. These three models are pre-trained generative unsupervised language models based on the Transformer architecture [@vaswani2017attention]. We started our experiment with BERT. Released by researchers at Google AI Language in 2018, the key innovation of BERT is its application of bidirectional training of the Transformer model, which enables BERT to perform text-infilling without fine-tuning.

@@

@@

@@
